---
title: "HarvardX Capstone My Own Project - Census Data"
author: "Frank Valdivia"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
    number_sections: yes
    df_print: paged
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
---

\newpage

# INTRODUCTION
 
 
The goal of this project is to build a classification model for the Census Data (1) downloaded from The UCI Machine Learning Repository.

According to the UCI website:

"Extraction was done by Barry Becker from the 1994 Census database.  A set of reasonably clean records was extracted using the following conditions: ((AAGE>16) && (AGI>100) && (AFNLWGT>1)&& (HRSWK>0))"


This dataset has 14 columns of which "Income" is the outcome. Income has two levels:

"<=50K" ">50K" 

The following is the list of the other features that the dataset has:

- age: continuous.

- workclass: Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked.

- fnlwgt: continuous.

- education: Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.

- education-num: continuous.

- marital-status: Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.

- occupation: Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct, Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces.

- relationship: Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried.

- race: White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black.

- sex: Female, Male.

- capital-gain: continuous.

- capital-loss: continuous.

- hours-per-week: continuous.

- native-country: United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands.
 
  
  
+++++++ IMPORTANT +++++++


To run this report, the file adult.zip should be downloaded from the following page and copied to the working directory of the R project running this report.

https://archive.ics.uci.edu/dataset/2/adult

Because processing time to generate the models was significant (more than one day), the  six models evaluated were generated and saved as RDS files in GitHub. Download the six models into your working directory:\

train_knn.rds: KNN model \
train_knn_cv.rds: KNN model with Cross Validation Model \
train_dt.rds: Decision Tree model \
train_dt_cp.rds: Decision Tree - Complexity Parameter model \
train_rf.rds: Random Forest model\
train_rf_2.rds: Random Forest wth Cross Validation model\


GitHub link:

https://github.com/frankvaldivia/Capstone-FV-DataScience/tree/main

+++++++++++++++++++++++++

 
 
Machine learning methods will be used, which means that two sub datasets will be created during this process:

A train set will be generated with the name: censusTrain; this dataset will be used to train the model.

A test set will be generated with the name:  censusTest; this dataset will be used to test the model trained using the train dataset.

The output will be a classification model and Accuracy will be used to measure quality of the model.

\newpage
# ANALYSIS AND METHODS

The overall process is as follows:

2.1. Loading libraries

2.2. Unzipping donloaded file and Generating datasets

2.3. Analysis and methods


```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

\newpage
## Loading libraries

To run the report the following R libraries need to be previously installed:

- library(tidyverse)
- library(caret)
- library(ggplot2)
- library(dplyr)
- library(lattice)
- library(gridExtra)
- library(beepr)
- library(dslabs)
- library(rpart)
- library(rpart.plot)
- library(randomForest)
- library(Rborist)
- library(class)
- library(grid)

Above libraries are loaded.

```{r loading_libraries, echo=FALSE}


if (!require("tidyverse")) install.packages("tidyverse")
if (!require("caret")) install.packages("caret")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("dplyr")) install.packages("dplyr")
if (!require("lattice")) install.packages("lattice")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("beepr")) install.packages("beepr")
if (!require("dslabs")) install.packages("dslabs")
if (!require("rpart")) install.packages("rpart")
if (!require("rpart.plot")) install.packages("rpart.plot")
if (!require("randomForest")) install.packages("randomForest")
if (!require("Rborist")) install.packages("Rborist")
if (!require("class")) install.packages("class")

library(tidyverse)
library(caret)
library(ggplot2)
library(dplyr)
library(lattice)
library(gridExtra)
library(beepr)
library(dslabs)
library(rpart)
library(rpart.plot)
library(randomForest)
library(Rborist)
library(class)
library(grid)

```

\newpage
## Unzipping donloaded file and Generating datasets


+++++++ IMPORTANT +++++++


To run this report, the file adult.zip should be downloaded from the following page and copied to the working directory of the R project running this report.

https://archive.ics.uci.edu/dataset/2/adult

Because processing time to generate the models was significant (more than one day), the  six models evaluated were generated and saved as RDS files in GitHub. Download the six files (models) into your working directory:\

train_knn.rds: KNN model \
train_knn_cv.rds: KNN model with Cross Validation Model \
train_dt.rds: Decision Tree model \
train_dt_cp.rds: Decision Tree - Complexity Parameter model \
train_rf.rds: Random Forest model\
train_rf_2.rds: Random Forest wth Cross Validation model\

GitHub link:\

https://github.com/frankvaldivia/Capstone-FV-DataScience/tree/main


+++++++++++++++++++++++++


```{r unzip_odwnload_load_datasets, echo=FALSE}

beep()

options(timeout = 120)

dl <- "adult.zip"

#  loading original train data file

adult_data_file <- "adult.data"
if(!file.exists(adult_data_file))
  unzip(dl, adult_data_file)

#  loading original test data file

adult_test_file <- "adult.test"
if(!file.exists(adult_test_file))
  unzip(dl, adult_test_file)

# both files will be put together to generate new train and test sets

# Loading into data frames and cleaning

# This step may take a few minutes

census_train <- as.data.frame(str_split(read_lines(adult_data_file), fixed(","), simplify = TRUE), stringsAsFactors = TRUE)

census_test <- as.data.frame(str_split(read_lines(adult_test_file), fixed(","), simplify = TRUE), stringsAsFactors = TRUE)


```

adult.zip has two files, The author already generated a train dataset and a test dataset. In this project, both files will be put together to create a whole Census_data dataset, which will be used to create a new train dataset and a new test dataset.

Following, 

1. Column names are assigned.
2. Empty rows are removed.
3. Temporary train and test data sets are created.

```{r organize_datasets, echo=TRUE}

beep()

# assigning column names to both temporary files

colnames(census_train) <- c("age", "workclass", "weight", "education", 
                            "educationlevel", "maritalstatus", "occupation", 
                            "relationship", "race", "sex", "capitalgain", 
                            "capitalloss", "hoursperweek", "nativecountry", 
                            "income")

colnames(census_test) <- c("age", "workclass", "weight", "education",
                           "educationlevel", "maritalstatus", "occupation",
                           "relationship", "race", "sex", "capitalgain",
                           "capitalloss", "hoursperweek", "nativecountry",
                           "income")

# removing empty rows

temp_train<-census_train[-c(32562,32563),]

temp_test<-census_test[-c(1,16283,16284),]


```

Temporary train and test data sets are created with the following structure and number of rows. These two data sets were originally created by the author and will be used to create a new master census data for this project.

```{r organize_datasets_show, echo=FALSE}

# Showing structure and number of records of the temp_train dataset

cat("\nTemporary train set: Columns\n")

colnames(temp_train) %>% knitr::kable()

cat("\nTemporary train set: Number of Records\n")

tibble(dim(temp_train))[1,1] %>% knitr::kable()

# Showing structure and number of records of the temp_test dataset

cat("\nTemporary test set: Columns\n")

colnames(temp_test) %>% knitr::kable()

cat("\nTemporary test set: Number of Records\n")

tibble(dim(temp_test))[1,1] %>% knitr::kable()

# generating the new master dataset

census_data <- rbind(temp_train,temp_test)



# Showing structure and number of records of the Census dataset

cat("\nNew Census data: Columns\n")

colnames(census_data) %>% knitr::kable()

cat("\nNew Census data: Number of Records\n")

tibble(dim(census_data))[1,1] %>% knitr::kable()

```

With the whole new Census Data set, predictors (columns) are redefined as Factors or Integers. The outcome (Y) was redefined as a string with two values:  “<=50K” “>50K”

Cleaning Census data continued with the removal of columns that will not be used in the analysis, empty records and the values for the outcome.

The census data set to be used in this project is as follows:

```{r organize_datasets_clean_data, echo=FALSE}


# converting to factors

census_data <- census_data %>%
  mutate(age = as.integer(age),
  	  workclass = factor(str_trim(as.character(workclass))),
      weight = as.integer(weight),
		  education = factor(str_trim(as.character(education))),
		  educationlevel = as.integer(educationlevel),
		  maritalstatus = factor(str_trim(as.character(maritalstatus))),
		  occupation = factor(str_trim(as.character(occupation))),
		  relationship = factor(str_trim(as.character(relationship))),
		  race = factor(str_trim(as.character(race))),
		  sex = factor(str_trim(as.character(sex))),
		  capitalgain = as.integer(capitalgain),
		  capitalloss = as.integer(capitalloss),
		  hoursperweek = as.integer(hoursperweek),
		  nativecountry = factor(str_trim(as.character(nativecountry))),
		  income = str_trim(as.character(income))
		  )
		
# cleaning up outcome variable: income
		
census_data <- census_data %>%
  mutate(income = gsub("<=50K.","<=50K",income))

census_data <- census_data %>%
  mutate(income = gsub(">50K.",">50K",income))

census_data <- census_data %>%
  mutate(income = factor(income))
  
# Removing variables that will not be included in the analysis

census_data = subset(census_data, select = -c(weight, educationlevel, capitalgain,capitalloss,nativecountry))

# Reordering categories for Education and Relationship

census_data$education <- factor(census_data$education, levels = c("Preschool", "1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th", "12th", "HS-grad", "Assoc-acdm", "Assoc-voc", "Some-college", "Bachelors", "Masters","Prof-school",  "Doctorate"))

census_data$relationship <- factor(census_data$relationship, levels = c("Husband", "Wife", "Unmarried", "Not-in-family", "Other-relative", "Own-child"))

# Showing structure and number of records of the census_data dataset

cat("\nClean Census data: Columns\n")

colnames(census_data) %>% knitr::kable()

cat("\nclean Census data: Head\n")

head(census_data) %>% knitr::kable()

cat("\nclean Census data: Number of Records\n")

tibble(dim(census_data))[1,1] %>% knitr::kable()

cat("\nclean Census data: levels for every predictor and outcome(Income)\n")

census_data %>% sapply(levels)


```

\newpage
## Analysis and methods



First, exploratory analysis will be carried out showing frequencies for predictors and the outcome (Income).

Then, a description of the measure of success (accuracy) will be included before elaborating on the models applied to Census data.

Every model will include specific results and plots and all will have their Accuracy estimated and compared.

\newpage
### Exploratory Analysis


Following are the frequencies for the following predictors. They are all categories:

- Work class
- Education
- Marital Status
- Race

```{r exploratory_analysis_1, echo=FALSE}

beep()

# Exploratory analysis



frequency_table <- table(census_data$workclass, census_data$income)

# Generation of frequencies for predictors

p1 <- as.data.frame(frequency_table)
colnames(p1) <- c("Category", "Income", "Frequency")

g1 <- ggplot(p1, aes(x=Category, y=Frequency, fill=Income)) + 
    geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none")  +
	ggtitle("Work class")

frequency_table <- table(census_data$education, census_data$income)

p2 <- as.data.frame(frequency_table)
colnames(p2) <- c("Category", "Income", "Frequency")

g2 <- ggplot(p2, aes(x=Category, y=Frequency, fill=Income)) + 
    geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none")  +
	ggtitle("Education")

frequency_table <- table(census_data$maritalstatus, census_data$income)

p3 <- as.data.frame(frequency_table)
colnames(p3) <- c("Category", "Income", "Frequency")

g3 <- ggplot(p3, aes(x=Category, y=Frequency, fill=Income)) + 
    geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none")  +
	ggtitle("Marital Status")

frequency_table <- table(census_data$race, census_data$income)

p4 <- as.data.frame(frequency_table)
colnames(p4) <- c("Category", "Income", "Frequency")

g4 <- ggplot(p4, aes(x=Category, y=Frequency, fill=Income)) + 
    geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  +
	ggtitle("Race")

grid.arrange(g1, g2, g3, g4, ncol = 2, 
               top="Frequencies", bottom="Census Data", 
               left=" ", right=" ")	
grid.rect(gp=gpar(fill=NA))

```

Previous plots suggest that there is some level of dependancy of income on some specific categories, i.e. HS-Grad, Bachelors, Some College in Education, or Private in Work class.


Following are the frequencies for the following predictors. They are all categories:

- Relationship
- Sex
- Occupation
- Income

```{r exploratory_analysis_2, echo=FALSE}


frequency_table <- table(census_data$relationship, census_data$income)

p5 <- as.data.frame(frequency_table)
colnames(p5) <- c("Category", "Income", "Frequency")

g5 <- ggplot(p5, aes(x=Category, y=Frequency, fill=Income)) + 
    geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none")  +
	ggtitle("Relationship")

frequency_table <- table(census_data$sex, census_data$income)

p6 <- as.data.frame(frequency_table)
colnames(p6) <- c("Category", "Income", "Frequency")

g6 <- ggplot(p6, aes(x=Category, y=Frequency, fill=Income)) + 
    geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  +
	ggtitle("Sex")

frequency_table <- table(census_data$occupation, census_data$income)

p7 <- as.data.frame(frequency_table)
colnames(p7) <- c("Category", "Income", "Frequency")

g7 <- ggplot(p7, aes(x=Category, y=Frequency, fill=Income)) + 
    geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none")  +
	ggtitle("Occupation")
	
frequency_table <- table(census_data$income)

p8 <- as.data.frame(frequency_table)
colnames(p8) <- c("Category", "Frequency")

g8 <- ggplot(p8, aes(x=Category, y=Frequency)) + 
    geom_bar(stat = "identity", fill ="darkgoldenrod3") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  +
	ggtitle("Income")

grid.arrange(g5, g6, g7, g8, ncol = 2, 
               top="Frequencies", bottom="Census Data", 
               left=" ", right=" ")	
grid.rect(gp=gpar(fill=NA))


```

Previous plots suggest that there is some level of dependancy of income on some specific categories, i.e. Female in Sex, or Own-child in Relationshp.
\newpage
The following graph shows the values of Income for the whole Census dataset. Because this is a two dimensional graph, only two predictors out of the nine are used, Education and Relationship.

Every dot shows the frequency of occurrences that have ">50K" as income in blue as a percentage of the total number of records for that combination of Education and Relationship. The size of the dot represents the number of records for that combination of Education and Relationship.


```{r Awesome_graph, echo=FALSE}


#  x = "Education", y = "Relationship"

df_income <- data.frame(census_data[, 3], census_data[, 6], census_data[,10])
colnames(df_income) <- c("Education", "Relationship", "Income")

# print(df_income)

proportion_df <- df_income %>%
  group_by(Relationship, Education) %>%
  summarise(
    count_filtered = sum(Income == ">50K"),
    total_count = n(),
    Red_to_Blue = count_filtered / total_count,
    .groups = "drop"
  )

# print(proportion_df)

base_income_plot <- ggplot(proportion_df, aes(x = Education, y = Relationship, 
                          color = Red_to_Blue, 
                          size = total_count)) +
    geom_point() +
    scale_color_gradient(low = "red", high = "blue") +
    labs(title = "Census Dataset\nOccurrences and degree of income\n Income<=50K (Red) Income>50K (Blue)",
         x = "Education", y = "Relationship",
         color = "% dots in Blue",
         size = "Occurrences") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) 
	
print(base_income_plot)


```

\newpage
### Accuracy and the Confusion Matrix


The accuracy of every model will be calculated and compared. Accuracy is the percentage of correct Income predictions on the Census data set. 

$$
Accuracy = \frac{RightPredictions }{TotalPredictions}
$$


The confusion Matrix will also be reported for every model.  The confusion Matrix from R summarizes the performance of the model. It compares the predicted values against the test values.

To run the models and get the confusin matrix, two sets were created from the Census data set. The train set and the test set, which is 30% of the whole Census data set.


```{r Create train and test datasets, echo=FALSE}

beep()

# Creating Train dataset and Test dataset. 

set.seed(42) 
test_index <- createDataPartition(census_data$income, times = 1, p = 0.7, list = FALSE) # create a 30% test set

censusTrain <- census_data[test_index,]
censusTest <- census_data[-test_index,]

```

Following are the structure and number of rows that each of these datasets has:



```{r show_datasets, echo=FALSE}

beep()

cat("\ncensusTrain: Head\n")

head(censusTrain) %>% knitr::kable()

cat("\ncensusTrain: Number of Records\n")

tibble(dim(censusTrain))[1,1] %>% knitr::kable()

cat("\ncensusTest: Head\n")

head(censusTest) %>% knitr::kable()

cat("\ncensusTest: Number of Records\n")

tibble(dim(censusTest))[1,1] %>% knitr::kable()

rm(temp_test, temp_train)

```

\newpage
### Models


The following models will be applied to the Census train data set and:

1. Predictions will be estimated
2. Plots will be generated
3. The confusion matrix will be produced comparing the predictions on the test set to the actual values of the test set.

Six models will be built and each model assessed against its Accuracy.

1. KNN (K-Nearest Neighbors): It classifies data points based on their similarity to existing data points. The similarity is measured thorugh the euclidian distance and the minimum distance is chosen to predict the values. During the running, this model was saved in a file (train_knn.rds) and uploaded to GitHub.
2. KNN with Cross Validation Model: It is a KNN model that uses tunning to find the optimum value of K. During the running, this model was saved in a file (train_knn_cv.rds) and uploaded to GitHub.
3. Decision Tree: It splits data into subsets based on a series of conditions identified based on minimizing a coefficient. Each internal node represents a test (Yes or No) on a preditor, each branch represents the outcome of the test, and each leaf node represents the predicted value "<=50K"  or  ">50K". During the running, this model was saved in a file (train_dt.rds) and uploaded to GitHub.
4. Decision Tree - Complexity Parameter: The complexity parameter optimizes the size and complexity of the tree during growth and can improve generalization and avoid overfitting.  During the running, this model was saved in a file (train_dt_cp.rds) and uploaded to GitHub.
5. Random Forest: It randomly generates many decision trees from the same data set to strenghten the prediction.  During the running, this model was saved in a file (train_rf.rds) and uploaded to GitHub.
6. Random Forest with Cross Validation: It applies cross validation to the Random Forst model.  During the running, this model was saved in a file (train_rf_2.rds) and uploaded to GitHub.

\newpage
#### First model: KNN\


The first model applied is KNN (K-Nearest Neighbors), which classifies data points based on their similarity to existing data points. The similarity is measured through the euclidian distance and the minimum distance is chosen to select the neighborhood and predict the values. 

During the running, this model was saved in a file (train_knn.rds) and uploaded to GitHub. The model file is saved on the working directory and load it every time to save running time.

The code to generate the model, save the model as a file and load the file with the model is as follows:



```{r KNN_model, echo=TRUE}

# KNN ====================================================================

beep()

# The code in comments might take from minutes to hours to run

Sys.time()		# Time is shown to track duration

set.seed(6)

# train_knn <- 	train(income ~ .,
#                     method = "knn",
#                     data = censusTrain,
#                     tuneGrid = data.frame(k = seq(3, 7, 2)))
# saveRDS(train_knn,"train_knn.rds")

# Loading from file to save running time
train_knn <- readRDS("train_knn.rds")

Sys.time()		# Time is shown to track duration

```


The following is the resulting KNN model. It shows that accuracy is maximum at K = 21, with an Accuracy of 0.8047201\



```{r KNN_model_report, echo=FALSE}

train_knn

# train_knn$bestTune
# train_knn$bestTune$k
# train_knn$finalModel

```

Following is the Variables Importance plot for the KNN model. This graph shows how much each predictor variable contributes to the model's predictive power.\

It shows that :\
- relationship\
- age\
- hoursperweek\

Are the most important variables for this model.

Next is a plot of Accuracy versus K showing that Accuracy is maximized at K = 21.



```{r KNN_model_VarImp, echo=FALSE}


plot(varImp(train_knn))

ggplot(train_knn, highlight = TRUE)


```


The following section shows the Confusion Matrix for this model. It shows:\

- Accuracy : 			0.81
- Sensitivity : 		0.9031
- Specificity : 		0.5140
- Balanced Accuracy :   0.7085


```{r KNN_model_accuracy, echo=FALSE}

beep()

# Generation of predictions using KNN

# This step may take a few minutes
# Sys.time()		# Time is shown to track duration
predicted_y_knn <- predict(train_knn, censusTest)
# Sys.time()		# Time is shown to track duration

Accuracy_knn <- confusionMatrix(predicted_y_knn, censusTest$income)$overall[["Accuracy"]]

confusionMatrix(predicted_y_knn, censusTest$income)

```

The First model and its Accuracy are as follows:


```{r KNN_model_results, echo=FALSE}

beep()

# The First model and its Accuracy are as follows:

Accuracy_results_t <- tibble(method = "1: KNN", Accuracy = Accuracy_knn, Parameter = paste("K = ",train_knn$bestTune$k))
Accuracy_results_t %>% knitr::kable()

```

\newpage
#### Second model: KNN with Cross Validation\

KNN with Cross Validation is a model that uses tunning to find the optimum value of K. 

During the running, this model was saved in a file (train_knn_cv.rds) and uploaded to GitHub. The model file is saved on the working directory and load it every time to save running time.

The code to generate the model, save the model as a file and load the file with the model is as follows:



```{r KNN_CV_model, echo=TRUE}

# KNN Cross Validation ==========================================

beep()

# The code in comments might take from minutes to hours to run

Sys.time()		# Time is shown to track duration

# control <- trainControl(method = "cv", number = 5, p = .9)
# train_knn_cv <- train(income ~ ., method = "knn",
#                       data = censusTrain,
#                       tuneGrid = data.frame(k = seq(3, 7, 2)),
#                       trControl = control)
# saveRDS(train_knn_cv,"train_knn_cv.rds")

# Loading model from file to save running time
train_knn_cv <- readRDS("train_knn_cv.rds")

Sys.time()		# Time is shown to track duration

```

The following is the resulting KNN_CV model. It shows that accuracy is maximum at K = 13\


```{r KNN_CV_report, echo=FALSE}

beep()

train_knn_cv

# train_knn_cv$bestTune
# train_knn_cv$bestTune$k
# train_knn_cv$finalModel

```

Following is the Variables Importance plot for the KNN_CV model. This graph shows how much each predictor variable contributes to the model's predictive power.\

It shows that :\
- relationship\
- age\
- hoursperweek\

Are the most important variables for this model.


```{r KNN_CV_model_VarImp, echo=FALSE}

# Variables Importance

plot(varImp(train_knn_cv))

ggplot(train_knn_cv, highlight = TRUE)


```
 
The previous plot shows Accuracy versus K showing that Accuracy is maximized at K = 13.

\newpage
The following plot is illustrative of the behavior of the KNN model. Because there are 9 predictors, they cannot be plotted in just one graph with two dimensions. A reduced KNN model was generated with just the two most important continuous variables for KNN:  "Age" and "Hours per Week". The contour of the reduced model's predictions was then plotted against CensusTrain. The contour shows the decisions' boundaries to assign "<=50K" or ">50K"


```{r KNN_CV_preview_plot, echo=FALSE}

beep()

# Generation of predictions usingn KNN CV

# Sys.time()		# Time is shown to track duration
predicted_y_knn_cv <- predict(train_knn_cv, censusTest)
# Sys.time()		# Time is shown to track duration

# dev.off()    # run before plotting KNN if issues with this graph - to clean plotting

# Plotting KNN most important continuous variables versus income

# Define the grid limits
x_min <- min(censusTest$age) - 1
x_max <- max(censusTest$age) + 1
y_min <- min(censusTest$hoursperweek) - 1
y_max <- max(censusTest$hoursperweek) + 1

# Create a grid of values
grid_income <- expand.grid(age = seq(x_min, x_max, by = 5),
                    hoursperweek = seq(y_min, y_max, by = 5))
					
k <- 13  # Number of neighbors
df_income <- data.frame(censusTrain[, 1], censusTrain[, 9])

predicted_income <- knn(train = df_income, test = grid_income, 
                         cl = censusTrain$income, k = k)
			
# Add predictions to the grid data
grid_income$PredictedIncome <- predicted_income

# Base plot with original data points
base_income_plot <- ggplot(censusTrain, aes(x = age, y = hoursperweek, 
                                     color = income)) +
  geom_point(size = 2) +
  labs(title = "Reduced k-NN(age, hoursperweek) Decisions' Boundaries",
       x = "Age",
       y = "Hours per Week") +
  theme_minimal()
  
# print(base_income_plot)

# Add decision boundaries using geom_contour
contour_income_plot <- base_income_plot +
  geom_contour(data = grid_income, aes(x = age, y = hoursperweek, 
                                z = as.numeric(PredictedIncome)),
               bins = 2, color = "black", linewidth=1) +
  scale_fill_manual(values = c("red", "green", "blue"))

# Display the plot
print(contour_income_plot)


```

The following section shows the Confusion Matrix for this model. It shows:\

- Accuracy : 			0.8101
- Sensitivity : 		0.8980
- Specificity : 		0.5308
- Balanced Accuracy :   0.7144

 


```{r KNN_CV_accuracy, echo=FALSE}


Accuracy_knn_cv <- confusionMatrix(predicted_y_knn_cv, censusTest$income)$overall[["Accuracy"]]
confusionMatrix(predicted_y_knn_cv, censusTest$income)


```

The second model and its Accuracy are as follows:


```{r KNN_CV_results, echo=FALSE}

beep()

# The second model and its Accuracy are as follows:

Accuracy_results_t <- bind_rows(Accuracy_results_t,
                          tibble(method="2: KNN with CV (Cross Validation)", 
						  Accuracy = Accuracy_knn_cv, 
						  Parameter = paste("K = ",train_knn_cv$bestTune$k)))
Accuracy_results_t %>% knitr::kable()


```

\newpage
#### Third model: Decision Tree\


Decision Tree splits data into subsets based on a series of conditions identified based on minimizing a coefficient. Each internal node represents a test (Yes or No) on a predictor, each branch represents the outcome of the test, and each leaf node represents the predicted value "<=50K"  or  ">50K". 

During the running, this model was saved in a file (train_dt.rds) and uploaded to GitHub. The model file is saved on the working directory and load it every time to save running time.

The code to generate the model, save the model as a file and load the file with the model is as follows:


```{r Decision_Tree_model, echo=TRUE}

# DECISION TREE ==========================================================

beep()

# The code in comments should run relatively fast in comparison to previous models

Sys.time()		# Time is shown to track duration

# train_dt <- rpart(income ~ ., data = censusTrain)
# saveRDS(train_dt,"train_dt.rds")

# Loading model from file to save running time
train_dt <- readRDS("train_dt.rds")

Sys.time()		# Time is shown to track duration

```
\newpage
Next is the plot of the Decision Tree model followed by the rules of the model and its Accuracy, which is maximum at 0.8187\


```{r Decision_Tree_report, echo=FALSE}

beep()


# display the decision tree

rpart.plot(x= train_dt, type= 5, extra = 0,tweak = 1.2)

# display model

train_dt

```

Following is the Variables Importance list for the Decision Tree model. This list shows how much each predictor variable contributes to the model's predictive power.\

It shows that :\
- relationship\
- maritalstatus\
- education\

Are the most important variables for this model.


```{r Decision_Tree_VarImp, echo=FALSE}

# Variables importance

listVarImp<-as.data.frame(varImp(train_dt))
listVarImp<-arrange(listVarImp, desc(Overall))
listVarImp %>% knitr::kable()

```


The following section shows the Confusion Matrix for this model. It shows:\

- Accuracy : 			0.8187
- Sensitivity : 		0.9402
- Specificity : 		0.4324
- Balanced Accuracy :   0.6863

 

```{r Decision_Tree_accuracy, echo=FALSE}

# Generation of predictions using Decision Tree

beep()

# Sys.time()		# Time is shown to track duration
predicted_y_dt <- predict(train_dt, censusTest, type = "class")
# Sys.time()		# Time is shown to track duration

Accuracy_dt <- confusionMatrix(predicted_y_dt, censusTest$income)$overall[["Accuracy"]]
confusionMatrix(predicted_y_dt, censusTest$income)


```

The third model and its Accuracy are as follows:


```{r Decision_Tree_results, echo=FALSE}

beep()

# The third model and its Accuracy are as follows:

Accuracy_results_t <- bind_rows(Accuracy_results_t,
                          tibble(method="3: Decision tree", 
						  Accuracy = Accuracy_dt, 
						  Parameter = paste(" ")))
Accuracy_results_t %>% knitr::kable()

```


\newpage
#### Fourth model: Decision Tree with Complexity Parameter\


The complexity parameter in a Decision Tree optimizes the size and complexity of the tree during growth and can improve generalization and avoid overfitting.

During the running, this model was saved in a file (train_dt_cp.rds) and uploaded to GitHub. The model file is saved on the working directory and load it every time to save running time.

The code to generate the model, save the model as a file and load the file with the model is as follows:


```{r DT_Complexity_Parameter_model, echo=TRUE}

# Decision Tree with Complexity Parameter (CP) =========================================

beep()

# The code in comments might take from minutes to hours to run

Sys.time()		# Time is shown to track duration

# train_dt_cp <- train(income ~ .,
#                     method = "rpart",
#                     tuneGrid = data.frame(cp = seq(0.0, 0.1, len = 25)),
#                     data = censusTrain)
# saveRDS(train_dt_cp,"train_dt_cp.rds")

# Loading model from file to save running time
train_dt_cp <- readRDS("train_dt_cp.rds")

Sys.time()		# Time is shown to track duration


```

Next is the Decision Tree model and the plot of Complexity Parameter versus Accuracy. It shows that Accuracy is maximum (0.8236853) for a CP of 0.004166667\




```{r DT_Complexity_Parameter_report, echo=FALSE}

beep()

train_dt_cp

plot(train_dt_cp)

```
\newpage
Following are the plot of the Decision Tree model with Complexity Parameter and the description of its rules.\

```{r DT_Complexity_Parameter_report_trees, echo=FALSE}

# Plotting decision tree

rpart.plot(x= train_dt_cp$finalModel)
# rpart.plot(x= train_dt_cp$finalModel, type= 5, extra = 0,tweak = 1.2)

# train_dt_cp$bestTune
# train_dt_cp$bestTune$cp

train_dt_cp$finalModel

```
\newpage
The Variables Importance plot is shown for the Decision Tree model. This plot shows how much each variable contributes to the model's predictive power.\


```{r DT_Complexity_Parameter_VarImp, echo=FALSE}

# Variables importance

# varImp(train_dt_cp$finalModel)
# varImp(train_dt_cp)
plot(varImp(train_dt_cp), top=15)


```


The following section shows the Confusion Matrix for this model. It shows:\

- Accuracy : 			0.8232
- Sensitivity : 		0.9351
- Specificity : 		0.4675
- Balanced Accuracy :   0.7013

 

```{r DT_Complexity_Parameter_accuracy, echo=FALSE}

# Generation of predictions using Decision Tree with Complexity Parameter

beep()

# Sys.time()		# Time is shown to track duration
predicted_y_dt_cp <- predict(train_dt_cp, censusTest)
# Sys.time()		# Time is shown to track duration

# compute accuracy
Accuracy_dt_cp <- confusionMatrix(factor(predicted_y_dt_cp), factor(censusTest$income))$overall["Accuracy"]
confusionMatrix(factor(predicted_y_dt_cp), factor(censusTest$income))

```

The fourth model and its Accuracy are as follows:


```{r DT_Complexity_Parameter_results, echo=FALSE}

beep()

# The fourth model and its Accuracy are as follows:

Accuracy_results_t <- bind_rows(Accuracy_results_t,
                          tibble(method="4: Decision tree with CP (Complexity Parameter)", 
						  Accuracy = Accuracy_dt_cp, 
						  Parameter = paste("cp = ", train_dt_cp$bestTune$cp)))
Accuracy_results_t %>% knitr::kable()

```


\newpage
#### Fifth model: Random Forest\


Random Forest generates many decision trees from the same data set to strenghten the prediction.  

During the running, this model was saved in a file (train_rf.rds) and uploaded to GitHub. The model file is saved on the working directory and load it every time to save running time.

The code to generate the model, save the model as a file and load the file with the model is as follows:


```{r Random_Forest_model, echo=TRUE}

# Random Forest =========================================

beep()

# The code in comments might take from minutes to hours to run

Sys.time()		# Time is shown to track duration

set.seed(6)
# For the generation of this report, the randomForest function was used
# train_rf<-randomForest(formula=income~.,
#                             data=censusTrain,
#                             ntree=500,
#							 do.trace=TRUE)
# saveRDS(train_rf,"train_rf.rds")

# Loading model from file to save running time
train_rf <- readRDS("train_rf.rds")

Sys.time()		# Time is shown to track duration

```

Next is the Random Forest model using 500 trees with 3 variables tried at each split.


```{r Random_Forest_report, echo=FALSE}

beep()

print(train_rf) # Print the model information


```

The variables Importance plot is shown for the Random Forest model. This plot shows how much each variable contributes to the models's predictive power. It shows that Age, Education and Relationship are the most important variables for the model.
This is also reflected in the plot of Variables versus MeanDecreaseGini that follows.




```{r Random_Forest_VarImp, echo=FALSE}


# Variables importance

listVarImp<-as.data.frame(varImp(train_rf))
listVarImp<-arrange(listVarImp, desc(Overall))
listVarImp %>% knitr::kable()

varImpPlot(train_rf)

plot(train_rf, log="y")


```

The previous plot shows how error is decreasing as number of trees is increasing
\newpage
The following two graphs show the dependance of the model on the values of the two most important variables: age and education.


```{r Random_Forest_VarImp_Partial, echo=FALSE}

# Importance of specific variables

partialPlot(train_rf, censusTrain, age, "<=50K")

partialPlot(train_rf, censusTrain, education, "<=50K", las = 3, xlab = "")

```

It shows that the model is more dependant for Ages less than 25 and Education as Doctorate and Masters.

The following section shows the Confusion Matrix for this model. It shows:

- Accuracy: 0.7497
- Sensitivity: 0.8271
- Specificity: 0.5037
- Balanced Accuracy: 0.8412


```{r Random_Forest_accuracy, echo=FALSE}

# Generation of predictions using Random Forest

beep()

# Sys.time()		# Time is shown to track duration
predicted_y_rf <- predict(train_rf, censusTest)
# Sys.time()		# Time is shown to track duration

Accuracy_rf <- confusionMatrix(factor(predicted_y_rf), factor(censusTest$income))$overall["Accuracy"]
confusionMatrix(factor(predicted_y_rf), factor(censusTest$income))

```

The fifth model and its Accuracy are as follows:


```{r Random_Forest_results, echo=FALSE}

beep()

# The fifth model and its Accuracy are as follows:

Accuracy_results_t <- bind_rows(Accuracy_results_t,
                          tibble(method="5: Random forest", 
						  Accuracy = Accuracy_rf, 
						  Parameter = paste("n of trees = 500, n of var = 3")))
Accuracy_results_t %>% knitr::kable()


```

\newpage
#### Sixth model: Random Forest with Cross Validation\


Cross validation is applied to the Random Forest model.  

During the running, this model was saved in a file (train_rf_2.rds) and uploaded to GitHub. The model file is saved on the working directory and load it every time to save running time.

The code to generate the model, save the model as a file and load the file with the model is as follows:



```{r Random_Forest_CV, echo=TRUE}

# Random Forest with CV =========================================

beep()

# The code in comments might take from minutes to hours to run

Sys.time()		# Time is shown to track duration

set.seed(6)

# train_rf_2 <- train(income ~ .,
#                     method = "Rborist",
#                     tuneGrid = data.frame(predFixed = 2, minNode = c(3, 50)),
#                     data = censusTrain)


# saveRDS(train_rf,"train_rf.rds")

# Loading model from file to save running time
train_rf_2 <- readRDS("train_rf_2.rds")

Sys.time()		# Time is shown to track duration

```

Next is the Random Forest model with Cross Validation generated using 300 trees with 3 variables tried at each split.


```{r Random_Forest_CV_report, echo=FALSE}

beep()

print(train_rf) # Print the model information



```

The Variables Importance plot is shown for the Random Forest CV model. This plot shows how much each variable contributes to the model's predictive power.

It shows that : 

- MaritalStatus-Married-Civ-Spouse
- Age
- MaritalStatus-Never-married

Are the most important variables for this model.


```{r Random_Forest_CV_VarImp, echo=FALSE}

# Variables importance
plot(varImp(train_rf_2), top = 15)


plot(train_rf_2, log="y")

```
The previous plot shows how Accuracy decreases as Minimal Node Size increases.

The following section shows the Confusion Matrix for this model. It shows:

- Accuracy : 0.816
- Sensitivity : 0.9775
- Specificity : 0.3026
- Balanced Accuracy : 0.6401


```{r Random_Forest_CV_accuracy, echo=FALSE}

# Generation of predictions using Random Forest with Cross Validation

beep()

# Sys.time()		# Time is shown to track duration
predicted_y_rf_CV <- predict(train_rf_2, censusTest)
# Sys.time()		# Time is shown to track duration

Accuracy_rf_CV <- confusionMatrix(factor(predicted_y_rf_CV), factor(censusTest$income))$overall["Accuracy"]
confusionMatrix(factor(predicted_y_rf_CV), factor(censusTest$income))

```



The sixth model and its Accuracy are as follows:


```{r Random_Forest_CV_results, echo=FALSE}

beep()

# The sixth model and its Accuracy are as follows:

Accuracy_results_t <- bind_rows(Accuracy_results_t,
                          tibble(method="6: Random forest with CV (Cross Valdidation)", 
						  Accuracy = Accuracy_rf_CV, 
						  Parameter = paste("")))
Accuracy_results_t %>% knitr::kable()


```


\newpage
# RESULTS


The following table shows all six models and their Accuracies:


```{r results_table, echo=FALSE}

beep()

Accuracy_results_t %>% knitr::kable()

```

Accuracy has been improving when incorporating Cross Validation or the Complexity Parameter to the base models.

The Decision Tree model including the analysis of the Complexity Parameter optimized Accuracy among the six models. Accuracy was maximized through iteration.

With the Decision Tree model with Complexity Parameter, an Income prediction model was built with an Accuracy of 82.32 percent.

\newpage
The following graph shows the actual values of Income for the  Census Test dataset. Because this is a two dimensional graph, the graph is illustrative and only two predictors out of the nine are shown, Education and Relationship.

Every dot shows the frequency of occurrences that have ">50K" as income in blue as a percentage of the total number of records for that combination of Education and Relationship. The size of the dot represents the number of records for that combination of Education and Relationship.


```{r results_awesome_graph, echo=FALSE}

#  x = "Education", y = "Relationship"

df_income <- data.frame(censusTest[, 3], censusTest[, 6], censusTest[,10])
colnames(df_income) <- c("Education", "Relationship", "Income")

# print(df_income)

# Summarize the count of value and total count per group
proportion_df <- df_income %>%
  group_by(Relationship, Education) %>%
  summarise(
    count_filtered = sum(Income == ">50K"),
    total_count = n(),
    Red_to_Blue = count_filtered / total_count,
    .groups = "drop"
  )

# print(proportion_df)

base_income_plot <- ggplot(proportion_df, aes(x = Education, y = Relationship, 
                          color = Red_to_Blue, 
                          size = total_count)) +
    geom_point() +
    scale_color_gradient(low = "red", high = "blue") +
    labs(title = "Census Test",
         color = "% dots in Blue",
         size = "Occurrences") +
  	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) 
#	theme(axis.text.x = element_blank(), legend.position = "none")

print(base_income_plot)

base_income_plot <- ggplot(proportion_df, aes(x = Education, y = Relationship, 
                          color = Red_to_Blue, 
                          size = total_count)) +
    geom_point() +
    scale_color_gradient(low = "red", high = "blue") +
    labs(title = "Census Test",
         x = "", y = "",
         color = "% dots in Blue",
         size = "Occurrences") +
  	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") 
#	theme(axis.text.x = element_blank(), legend.position = "none")

# print(base_income_plot)

beep()

```
\newpage
The following graph shows an illustrative visual comparison between the actual values of the census test dataset and a graph generated with the predicted values of the model with the hightest Accuracy (Decision Tree with Complexity Parameter).


```{r results_awesome_graph_dt_cp, echo=FALSE}

#  x = "Education", y = "Relationship"

df_income <- data.frame(censusTest[, 3], censusTest[, 6], predicted_y_dt_cp)
colnames(df_income) <- c("Education", "Relationship", "Income")

# Summarize the count of value and total count per group
proportion_df <- df_income %>%
  group_by(Relationship, Education) %>%
  summarise(
    count_filtered = sum(Income == ">50K"),
    total_count = n(),
    Red_to_Blue = count_filtered / total_count,
    .groups = "drop"
  )

predicted_y_dt_cp_plot <- ggplot(proportion_df, aes(x = Education, y = Relationship, 
                          color = Red_to_Blue, 
                          size = total_count)) +
    geom_point() +
    scale_color_gradient(low = "red", high = "blue") +
    labs(title = "Decision Tree with CP",
         x = "", y = "",
         color = "% dots in Blue",
         size = "Occurrences") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") 
#	theme(axis.text.x = element_blank(), legend.position = "none") 

# print(predicted_y_dt_cp_plot)

beep()

grid.arrange(	
				base_income_plot, 
				predicted_y_dt_cp_plot, 
				ncol = 2, 
				top="Census Test values and Decicion Tree CP predictions on Income", bottom="", 
				left="", right="")	
				
grid.rect(gp=gpar(fill=NA))


```


\newpage
The following graph shows an illustrative visual comparison between the actual values of the census test dataset and a graph generated with the predicted values of the model with the second hightest Accuracy (Decision Tree).



```{r results_awesome_graph_dt, echo=FALSE}

#  x = "Education", y = "Relationship"

df_income <- data.frame(censusTest[, 3], censusTest[, 6], predicted_y_dt)
colnames(df_income) <- c("Education", "Relationship", "Income")

# Summarize the count of value and total count per group
proportion_df <- df_income %>%
  group_by(Relationship, Education) %>%
  summarise(
    count_filtered = sum(Income == ">50K"),
    total_count = n(),
    Red_to_Blue = count_filtered / total_count,
    .groups = "drop"
  )

predicted_y_dt_plot <- ggplot(proportion_df, aes(x = Education, y = Relationship, 
                          color = Red_to_Blue, 
                          size = total_count)) +
    geom_point() +
    scale_color_gradient(low = "red", high = "blue") +
    labs(title = "Decision Tree",
         x = "", y = "",
         color = "% dots in Blue",
         size = "Occurrences") +
	theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") 
	
# print(predicted_y_dt_plot)

beep()


grid.arrange(	
				base_income_plot, 
				predicted_y_dt_plot, 
				ncol = 2, 
				top="Census Test values and Decicion Tree predictions on Income", bottom="", 
				left="", right="")	
				
grid.rect(gp=gpar(fill=NA))


```



\newpage
# CONCLUSION


In this analysis, six models were considered and accuracy was used to measure the predictive power of each model.


The models were increasing accuracy to get to the maximum when the analysis of Complexity Parameter was applied to the Decision Tree model.

Using Complexity Parameter, Accuracy was maximized through iteration.

With the Decision Tree model with Complexity Parameter, an Income prediction model was built with an Accuracy of 82.32 percent.

One potential limitation of the analysis is that the dataset has few records with income >50K in comparison to records with Income <=50K. This maybe a problem of Prevalence (when the ratio of Trues is close to 0 or 1).

A couple of ways to address the issue of prevalence are:

- generating a sample without prevalence
- analyzing sensitivity and specificity


An additional step that may improve the prediction is to ensemble all six models into one. Ensembles combine multiple machine learning algorithms into one model to improve predictions (2).

\newpage
# ADDITIONAL GRAPHS


```{r additional_graphs, echo=FALSE}

beep()


base_income_plot <- ggplot(censusTest, aes(x = education, y = relationship, color = income)) +
  geom_point(shape= 15, size = 6) +
  labs(title = "Sample of Census Test Dataset on Income\nRed: <=50K   Blue: >50K",
       x = "Education",
       y = "Relationship") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))



predicted_income_plot_KNN <- ggplot(censusTest, aes(x = education, y = relationship, color = predicted_y_knn)) +
	geom_point(shape= 15, size = 6) +
    labs(title = paste("k-NN \n(Accuracy=",round(Accuracy_knn,2),")"),
       x = "Education",
       y = "Relationship") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))



predicted_income_plot_KNN_CV <- ggplot(censusTest, aes(x = education, y = relationship, color = predicted_y_knn_cv)) +
	geom_point(shape= 15, size = 6) +
    labs(title = paste("k-NN Cross Validation\n (Accuracy=",round(Accuracy_knn_cv,2),")"),
       x = "Education",
       y = "Relationship") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))



predicted_income_plot_dt <- ggplot(censusTest, aes(x = education, y = relationship, color = predicted_y_dt)) +
	geom_point(shape= 15, size = 6) +
    labs(title = paste("Decision Tree \n(Accuracy=",round(Accuracy_dt,2),")"),
       x = "Education",
       y = "Relationship") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))



predicted_income_plot_dt_cp <- ggplot(censusTest, aes(x = education, y = relationship, color = predicted_y_dt_cp)) +
	geom_point(shape= 15, size = 6) +
    labs(title = paste("Decision Tree\nComplexity Parameter\n (Accuracy=",round(Accuracy_dt_cp,2),")"),
       x = "Education",
       y = "Relationship") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))



predicted_income_plot_rf <- ggplot(censusTest, aes(x = education, y = relationship, color = predicted_y_rf)) +
	geom_point(shape= 15, size = 6) +
    labs(title = paste("Random Forest \n\n(Accuracy=",round(Accuracy_rf,2),")"),
       x = "Education",
       y = "Relationship") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))


predicted_income_plot_rf_CV <- ggplot(censusTest, aes(x = education, y = relationship, color = predicted_y_rf_CV)) +
	geom_point(shape= 15, size = 6) +
    labs(title = paste("Random Forest with CV\n\n(Accuracy=",round(Accuracy_rf_CV,2),")"),
       x = "Education",
       y = "Relationship") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1))


grid.arrange(	
				base_income_plot, 
				ncol = 1, 
				top="Test data on Income ", bottom="", 
				left=" ", right=" ")	
				
grid.rect(gp=gpar(fill=NA))


```

\newpage

```{r additional_graphs2, echo=FALSE}


grid.arrange(	
				predicted_income_plot_KNN, 
				predicted_income_plot_KNN_CV, 
				ncol = 2, 
				top="Sample of model predictions on Income", bottom="Models", 
				left=" ", right=" ")	
				
grid.rect(gp=gpar(fill=NA))



grid.arrange(
				predicted_income_plot_dt,
				predicted_income_plot_dt_cp,
				ncol = 2, 
				top="Sample of model predictions on Income", bottom="Models", 
				left=" ", right=" ")	
				
grid.rect(gp=gpar(fill=NA))


grid.arrange(
				predicted_income_plot_rf,
				predicted_income_plot_rf_CV,
				ncol = 2, 
				top="Sample of model predictions on Income", bottom="Models", 
				left=" ", right=" ")
				
grid.rect(gp=gpar(fill=NA))

```

\newpage
# REFERENCES


 1. Becker, B. & Kohavi, R. (1996). Adult [Dataset]. UCI Machine Learning Repository. https://doi.org/10.24432/C5XW20.
 
 2.  Irizarry, R. Introduction to Data Science  - Data Analysis and Prediction Algorithms with R (32.5 Ensembles)
 https://rafalab.dfci.harvard.edu/dsbook/machine-learning-in-practice.html#ensembles
 
 


